A method for inference of fundamental stellar parameters using high resolution spectra

### 2014-01-15 paper/paper.tex
Running lkca15 on kurucz

%\documentclass[manuscript]{aastex} %one-column, double-spaced document
\documentclass[preprint]{aastex} %double-column, single-spaced document:
%\documentclass[iop,floatfix]{emulateapj}

@@

\newcommand{\prob}{{\rm prob}}
\newcommand{\qN}{\{q_i\}_{i=1}^N}
\newcommand{\yN}{\{y_i\}_{i=1}^N}
\newcommand{\vt}{\vec{\theta}}
\newcommand{\vstar}{\vt_{\star}}
\newcommand{\vN}{\vt_{\rm N}}
\newcommand{\vc}{\vec{c}}
\newcommand{\fM}{f_{\rm M}}
\newcommand{\fD}{f_{\rm D}}
\newcommand{\vD}{\vec{D}}
\newcommand{\dd}{\,{\rm d}}
\newcommand{\trans}{\mathsf{T}}

@@

\title{A Bayesian method for inference of fundamental stellar parameters}
\author{\today{}\\
\medskip
Ian~Czekala\altaffilmark{1} et al.
%Author2\altaffilmark{2},
}

@@

%\begin{abstract}
%\end{abstract}
%\keywords{}

@@

The fundamental parameters of stars, such as mass, luminosity, effective photospheric temperature, surface gravity, and metallicity are of supreme interest in astrophysics, for a wide variety of reasons. Exoplanet properties are found relative to the properties of the host star, and have important implications for models of planet formation. The properties are also important to the study of disks, for the incident radiation field sets photoevaporation, influences disk dispersal. 
Many methods exist for determination of stellar parameters: SPC \citep{blj+12}, MOOG, SME \citep{vp96}, and these are robust for FKG stars, where there is a well-defined continuum and the stars are ``normal.'' However, for many young T-Tauri stars, the spectrum is contaminated by accretion, veiling, and fractional star spot coverage, which can be a nuisance and cause systematic errors if traditional techniques are used.

\section{The model}

To summarize our model, we use Bayes rule to design a posterior probability function
\begin{equation}
  \overbrace{p(\vt | \vD)}^{\rm posterior} \propto \underbrace{p(\vD | \vt)}_{\rm likelihood} \; \overbrace{p(\vt)}^{\rm prior}
  \label{eqn:posterior}
\end{equation}
where the parameter vector is comprised of both stellar parameters and calibration parameters $\vt = \{\vstar, \vN \}$. Stellar parameters include $\vstar = \{T_{\rm eff}, \log(g), [{\rm Fe}/{\rm H}], v \sin i, v_z, A_v, R_\star^2/d^2 \}$ while the data vector $\vD = \fD(\lambda)$ represents a high resolution spectrum. For demonstration purposes in this paper, we use a 51-order echelle spectrum from the \emph{TRES} spectrograph at a resolution of $R=48,000$ ($6.8\;{\rm km/s}$), which spans the full optical range ($3800-9000$\AA), though our technique is applicable to any high resolution optical/near-IR spectrum.

@@

Because matrix multiplication is associative, we can sum ${\bm W(\lambda)}$ and $\vec{T}^\trans(\lambda)$ across all $\lambda$, and define
\begin{align}
  {\bm A} &= c_0^2 \sum_{\lambda} \frac{\fM^2(\lambda)}{\sigma^2(\lambda)} {\bm W}(\lambda) \\
  \vec{B} &= \sum_{\lambda} \frac{-\fM^2(\lambda) c_0^2 + \fD(\lambda) \fM (\lambda) c_0}{\sigma^2(\lambda)} \vec{T}(\lambda)\\
  g &= -\frac{1}{2} \sum_{\lambda} \frac{c_0^2 \fM^2(\lambda) - 2 c_0 \fD(\lambda) \fM(\lambda) + \fD^2(\lambda)}{\sigma^2(\lambda)} 
\end{align}
Where ${\bm A}$, $\vec{B}$, and $g$ are each a function of $\vstar$, $\vN$, and $\vD$. Now we can rewrite Equation~\ref{eqn:lnprob2} as 
\begin{align}
  \ln \bigl [p(\vstar, \vN | \vD) \bigr] &\propto - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans \vc + g\\
  p(\vt | \vD) = p(\vstar, \vN | \vD) &\propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans \vc + g \right )
  \label{eqn:lnprob_matrix}
\end{align}
Here we have the full posterior probability distribution defined in Equation~\ref{eqn:posterior}, assuming flat priors. This posterior is a function of both the stellar parameters $\vstar$ and the nuisance parameters $\vN$. Generally, we are most interested in the stellar parameters after they have been marginalized over the nuisance parameters. This marginalization encapsulates the uncertainty due to the inference on the nuisance parameters. Because this probability function is a multi-dimensional Gaussian in the higher order Chebyshev coefficients $\vc = \{c_1, c_2, \ldots, c_N \}$, we can do this marginalization analytically, and then sample from the marginalized distribution. 
\begin{equation}
  p(\vstar, c_0 | \vD) = \int p(\vstar, c_0, \vc\; | \vD) \dd \vc
\end{equation}
The analytic multi-dimensional Gaussian integral with linear term \citep{sgd+09} yields 
\begin{align}
  p(\vstar, c_0 | \vD) &\propto \int \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans \vc + g \right ) \dd \vc\\
  p(\vstar, c_0 | \vD) &\propto \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A} \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^\trans {\bm A}^{-1} \vec{B} + g \right )}
\end{align}
Where $N$ is the number of Chebyshev coefficients marginalized over, in our case $N=3$. By doing this marginalization we have reduced the dimensionality of our posterior space from $200+$ parameters ($\sim 7$ stellar parameters plus 3 Chebyshev coefficients for each of 51 echelle orders) to $\sim 60$ ($\sim 7$ stellar parameters and one $c_0$ for each of the 51 echelle orders). This enables a dramatic speedup in the MCMC sampling. For flux-calibrated spectra where the variation in $c_0$ is expected to be small (less than XX\%), we can approximate $k$ in a different manner which allows marginalization over the $c_0$ coefficients as well, reducing the dimensionality of the space to only the $\sim 7$ stellar parameters. This approach is detailed in \S\ref{sec:gaussian_simplification}.

>>>>

\documentclass[preprint]{aastex} %double-column, single-spaced document:
%\documentclass[iop,floatfix]{emulateapj}

@@

\newcommand{\prob}{{\rm prob}}
\newcommand{\qN}{\{q_i\}_{i=1}^N}
\newcommand{\yN}{\{y_i\}_{i=1}^N}
\newcommand{\vt}{\vec{\theta}}
\newcommand{\vg}{\vt_{\star, {\rm grid}}}
\newcommand{\vstar}{\vt_{\star}}
\newcommand{\vN}{\vt_{\rm N}}
\newcommand{\vc}{\vec{c}}
\newcommand{\fM}{f_{\rm M}}
\newcommand{\fD}{f_{\rm D}}
\newcommand{\vD}{\vec{D}}
\newcommand{\dd}{\,{\rm d}}
\newcommand{\trans}{\mathsf{T}}
\newcommand{\Z}{[{\rm Fe}/{\rm H}]}
\newcommand{\A}{[\alpha/{\rm Fe}]}

@@

\title{A method for inference of fundamental stellar parameters using high resolution spectra}
\author{\today{}\\
\medskip
Ian~Czekala\altaffilmark{1} et al.
%Author2\altaffilmark{2},
}

@@

\begin{abstract}
  Matching observed spectra of stars against a grid synthetic spectra is a popular approach to determine fundamental stellar parameters such as effective temperature, surface gravity, and metallicity. However, a fundamental problem with this approach is that the synthetic spectra contain systematic errors such as wrong or missing spectral lines. In the presence of such errors, a simple $\chi^2$ fit between spectra will likely yield biased parameters and incorrect uncertainty estimates. This behaviour is analogous to the way a few strong outliers might severely bias the fit to a straight line. These spectral outliers are difficult to identify \emph{a priori}; procedures such as sigma clipping might be prone to their own systematic effects and discard useful information. We outline an approach to probabilistically identify and account for spectral features that deviate from reality. Such an approach is most needed when operating at high spectral resolution ($R >$ 10,000), since this is when systematic deviations begin to appear most striking, however our technique is still valid for lower resolution spectra as well. We also outline some approaches for spectral fits to brown dwarfs and planets, where instead of spectral lines there may be systematic deviations of a larger swath of the spectrum.
\end{abstract}

%\keywords{}

@@

The fundamental parameters of stars---mass, luminosity, effective photospheric temperature, surface gravity, and metallicity--are of supreme interest to astrophysicists for a wide variety of reasons. A salient example is that most newly discovered exoplanets must necessarily their mass and radius measured relative to the properties of their host stars (e.g. \citealt{tfs+12,blj+12,ssm+13}). The properties of stars are of course also important in their own right because they inform us about our ideas of stellar evolution \citep{dm97, bca+02}. Because stars are the fundamental units of the cosmos, techniques to infer the properties of stars date back to the dawn of astronomy. We briefly review some of the modern and popular techniques in the literature. Approaches can be broadly classified into techniques that compare stars by some broadly defined indices and techniques that attempt to forward-model an observed stellar spectrum by doing a pixel-to-pixel comparison.

\subsection{Similarity techniques}
Many of existing techniques were written to analyze low to medium resolution ($R < $ 10,000) spectra. I am calling them similarity techniques because they compare characteristics of spectra, not actual pixels.

\paragraph{Line indexes} By cataloguing the equivalent widths of chosen spectral lines into indexes, one can correlate the index with stellar properties and identify features that are sensitive temperature, surface gravity, or metallicity. Then, one matches the indecies of a new star against these relationships and derives a set of stellar parameters. A recent example using M dwarfs is \citet{nci+14}. This technique works, but there are many things that could be improved. First, the selection of which lines to measure is arduous and based upon expert knowledge of previously catalogued stars. This is both a disadvantage and an advantage. One one hand by choosing a subset of lines to fit, one may be unwittingly throwing away useful information. On the other hand, it allows the modeller to discard regions of the spectrum that do not follow a simple trend with stellar parameters. Second, while measuring equivalent widths is easy to do, this discards any useful information contained in the shape of the line profile.

\paragraph{Principal component analysis} PCA reduces the dimensionality of the data space and searches for the most salient features of the spectra \citep{ptb14}. How does one properly estimate parameter uncertainties in such an approach? It seems like this approach exists because it is an easy thing to do.

Matrix inversions: MATISSE \citep{rbd06}. This method is similar to PCA in that it derives basis vectors that describe the fundamental parameters of the star. In the PCA analysis, the principal components do not necessarily correspond to true physical quantities. The MATISSE algorithm determines weights that ensure the components are similar to the stellar parameters. They only test the algorithm against input synthetic stellar spectra to detect biases in the parameters, and against the Sun and Arcturus. They do acknowledge that there will be systematic deviations between the synthetic spectra and real spectra, but they do not account for the possible inflation of error estimates.

Given the inherent systematic uncertainties involved in each method, some groups elect to combine the inference from different algorithms, such as (SLOAN SSPP) \citep{lbs+08}. This approach includes machine learning along with many of the previously mentioned techniques.

\subsection{Pixel to pixel techniques using spectral libraries}
Many high quality synthetic spectral grids are now available. Nearly all synthetic stellar libraries are parameterized by the ``fundamental'' stellar parameters of effective temperature, surface gravity, metallicity, and alpha enhancement. We define this combination as 
\begin{equation}
  \vg = \{T_{\rm eff}, \log(g), \Z, \A \}
\end{equation}
Naively, it makes sense to compare an observed spectrum against spectra from the synthetic grid. Using some goodness of fit statistic one can identify the $\vg$ that best reproduce the observed spectrum.

\paragraph{SPC} A very popular method behind the Kepler follow up program is Stellar Parameter Classification (SPC) \citep{blj+12}). This compares high resolution optical echelle spectra ($R \sim$ 45,000) to high resolution synthetic grids computed with Kurucz models by a cross-correlation method. The best fit parameters and uncertainties are calculated by the relative values of the cross-correlation coefficient. It is uncertain how statistically justified this uncertainty estimate is. While this is a pixel-to-pixel method, it uses only a small region of the optical spectrum ($5000 - 5300$\AA) where the spectral library has been carefully tweaked\footnote{changes to line centers and $\log(gf)$ values were on the table} to match  high resolution models of the Sun and Vega.

Most other pixel-to-pixel techniques involve comparison via a $\chi^2$ likelihood function.

\paragraph{ULySS} ULySS \citep{kpb+09} is a mature package to make comparisons for individual spectra as well as unresolved stellar populations. Any spectral library can be used as a backend, including empirical spectral libraries, such as the one from ELODIE \citep{psk+07}. population uses an iterative kappa-sigma cleaning method to discard ``bad'' lines. They test for convergence to the same answer by starting guesses located at different parameters and seeing if the same lines are masked.

Takes into account many of the things that need to be considered, including multiplication by a polynomial and rejection of bad data or model points.

Bayesian methods exist, but they use a $\chi^2$ likelihood function \citep{sdm+07}.

\citet{sb13} mention a comprehensive Bayesian method that can fit low or high resolution spectra. They too, mention that they use a $\chi^2$ likelihood. They allude to using special masks for the spectra that block noisy and uninformative regions of the spectrum, however they do not go into further detail about how these masks were identified. Also, some stellar grids focus on the most important parts of the spectrum, as identified by experts. Quoted from \citet{sb13}, grids ``sample the wavelength windows around the spectral features important for diagnostic of FGKM stars: 3850 - 4050\AA\ (Ca I lines), 4350 - 4450\AA\ (G-band, CN sensitive), 4600 - 4900\AA\ (H$\beta$), 5100 - 5300\AA\ (Mg I triplet, main gravity diagnostics), 6400 - 6640\AA (H$\alpha$), 8400 - 8800\AA\ (Ca II triplet, also used in \emph{Gaia} and in RAVE stellar survey).'' These areas can use intense scrutiny from a spectral modeller to guarantee that the models match reality.

\citet{mga13} fit BT-Settl PHOENIX models to M dwarfs. Surely these models have some serious problems with them. They use a $\chi^2$-like likelihood function, and iterate to clean bad pixels. They iterate the method against their grid of $\sim 20$ M dwarfs and track which pixels consistently have the high residuals (median residual amount $> 10$\%, or ten times their measurement error) across all stars. Presumably, this would be due to a bad opacity for a certain line or molecular feature. Then, they mask these pixels and repeat the fitting. They claim that after two iterations the masks do not change. They note that rather than selecting good and bad pixels with a binary mask, ``a more nuanced weighting scheme would be to weight each interval (pixel) according to how consistent it is with the models, or test different weighting schemes to see which gives the best agreement with the bolometric temperatures. However regions with modest agreement between the real and synthetic spectra may contain more temperature information than regions with slightly better matches, and we have no a priori information about what spectral regions are the most temperature sensitive.''

\subsection{Line fitting using spectral synthesis}
This is an improvement over the line index method, since now one includes the information contained in the line profile. With quality stellar spectral models, one can identify a few well-modelled lines to fit. By choosing a few good lines to fit, one could determine stellar parameters by matching line profiles for a few select lines \citep{dj03}. However, this is still subject to the need to identify promising regions of the spectrum.

MOOG\footnote{\url{http://www.as.utexas.edu/~chris/moog.html}} \citep{sne73}, SME\footnote{\url{http://www.stsci.edu/~valenti/sme.html}} \citep{vp96}, actually synthesize line profiles on the fly, compared to using a library of spectral models. This has the advantage of enabling individual tweaks to elemental abundances. And interpolation in the model atmosphere space, rather than the spectral line space. However, for many young T-Tauri stars, the spectrum is contaminated by accretion, veiling, and fractional star spot coverage, which can be a nuisance and cause systematic errors if traditional techniques are used. Matching equivalent widths. Interestingly, MOOG sythesizes a spectrum on the fly.

SME allows adjustment of oscillator strengths. We could download the synthesis code for SME and synthesize our own spectra on the fly. The biggest impediment here would be that the simplest atmospheres will not get certain lines correct. The powerful synthesis codes (PHOENIX) will get the atmospheres and lines correct, but they take a long time to run (~hours) per model, and the code is not publicly available. To do a full-spectrum fit for these sorts of codes would likely be too computationally difficult. However, it would allow access to fitting in the parameter space that we want.

\subsection{Stellar parameter estimation in a utopia}

Although the spectral synthesis domain is the correct domain to solve for stellar parameters, the radiation transfer is too slow.

The fitting difficulties of the previous section are a result of trying to fit data to a model in a smaller parameter space. Yes, the combination of $\vg$ is an excellent proxy to describe a star, but in truth (and at high resolution) each star is slightly different. If we think back to first principles, the ``model'' that is actually producing the synthetic spectrum has many, many parameters that have been collapsed down into $\vg$. These parameters include the atmospheric structure (relations of temperature and pressure as a function of stellar depth), individual elemental abundances (all models quote some value relative to solar, a fraction of models allow changes in alpha abundances), and atomic constants describing the opacity contribution of a line. \emph{All} of these parameters have uncertainties in them, some of these are rather dramatic (one continually hears lamentation about the opacities and line lists\dots), yet for any given stellar model these parameters are \emph{fixed}. There may yet be unconsidered stellar physics necessary to explain a small fraction of lines (say for example the cores of Ca lines), but the field of stellar modelling has progressed to the state that I would have the confidence to say that if a group focused all of their attention on a single line, and using only their current physical models and making tweaks to the afore-mentioned lines, they could produce a satisfactory model.

In reality, this has been done using the Kurucz models in an unpublished grid of stellar spectra, over the range $5000 - 5300$ \AA.

The real problem is much more rich, as is the amount of information one could extract from a high resolution optical spectrum. One would want to fit for individual abundances, marginalize over oscillator strengths.

The unfortunate reality is that sampling in the stellar model space for the full optical spectrum is not yet a computational reality. There do exist codes (MOOG) that can do this approach but only for a small ($\sim$ a few lines) at once, and using simplifications. Through collaboration with a stellar atmospheric expert, such an approach may soon be a possibility. If the computation of a given model could be reduced to less than a minute, efficient MCMC sampling parallelized on a cluster (using \texttt{emcee}, for example). However, as it stands without a collaborator, I do not have the expertise nor the time to pursue such an approach. Therefore, we are reduced to fitting in the space of $\vg$ and devising ways to deal with the issues.

\section{General problems with the spectral library approach}

Pixel-to-pixel approaches have the potential to be much more powerful than other historical stellar parameter estimation techniques. When comparing directly in the data space, we have the ability to add in other physics as we see fit.

because we can fit multiple stars at once (say for unresolved binary or multiple systems), fit for velocity components, etc. Many of the most interesting systems for which we would want to determine stellar parameters are in multiple systems anyway, since we can dynamically determine masses for these systems and thus compare to evolutionary models. Determine veiling. And a perhaps wooden-headed belief that using the full spectrum \emph{should} give us a more accurate and less-biased posterior estimate. Such a belief may be foolish and in reality different parts of the spectrum could give completely different parameter estimates.

Many, many methods exist for comparison of stellar spectra, including many ``Bayesian'' approaches in the pixel-to-pixel space. The nature of the Bayesian-ness, however, is usually limited to priors on temperature, surface gravity, and metallicity. These approaches do not explicitly state the likelihood function that they use, or if they do, it is a $\chi^2$ function.

The more successful of these models use low or moderate resolution spectra ($R \sim$ 1,000 to 10,000 ish). Or, a grid over a narrow range of wavelength space with finely tuned atomic constants.

\textbf{The true fundamental problem with this approach is that the stellar spectra are not a perfect representation of reality}. Assumptions about Poisson errors break down, and fits become contaminated by systematics. There are types of errors which occur.

FIGURE: 4 Panel Figure, from left to right, showing Class 0 through III errors. Use Mg b lines, other big line near Mg b triplet, order 24 line, and then Fe I/II line complex that doesn't fit properly.

\paragraph{Class I} A spectral line appears in the data but not in the model spectrum over a wide range of probable stellar parameters (1000 K in $T_{\rm eff}$, and 1 dex in $\log(g)$, $\Z$, and $\A$. Such a type of error is perhaps the easiest to understand and to forgive. For example, such a line could easily appear in the data due to interstellar absorption or telluric absorption, something which a stellar model could never produce. The model could also fail to reproduce a true stellar line due to a missing opacity source.

\paragraph{Class II} A spectral line appears in the model but not in the data over a wide range of probable stellar parameters (1000 K in $T_{\rm eff}$, and 1 dex in $\log(g)$, $\Z$, and $\A$. This is a more puzzling error, because presumably atomic constants would be much more likely to be \emph{missing} than added in by accident. Presumably this is a product of bad atmospheric structure (temperature or pressure is wrong at some point?) that excites a line when it shouldn't otherwise.

\paragraph{Traditional ways to model Class I and Class II errors} are a form of sigma clipping. Mann et al. compares the stars to models over a wide range of parameters. One can track how these lines might be bad because Class I residuals (data - model)/sigma will always be negative (since we are dealing with absorption lines only) and not change significantly with $\vg$, since we the model can only produce continuum. Class II residuals will always be positive and should decrease with decreasing metallicity, since lower metallicity will decrease the strength of all lines.

If a Class I or II error occurs on a line, that region of the spectrum is relatively useless for determining spectral parameters. Authors like Mann et al. or Kholeva et al. find these regions and simply mask them out of the final computation. Kholeva et al., and others approach in an iterative manner. Kholeva et al. approaches are more sophisticated because they track convergence. This is a sigma clipping approach, and while we could go and model these a la Hogg 2010, the approach is not super justified, result is probably not too wrong (similar to actual sigma clipping). Given that Kholeva et al. track convergence maps, I would place a higher degree of confidence in the answer.

In our later approaches in Sec XX, how might we put a prior on $\Delta \vg$, the range of parameters over which the residuals will be correlated? Maybe we could Gibbs sample in the same way as the bad flags. We'd have to keep a bunch of spectra with varying $\vg$ in memory to make this comparison easy.

\paragraph{Class III} In truth, this is just a superset of both Class I and Class II. Class III errors are when the line strengths between the data and model do not match at a $\vg$ that fits a majority of lines. What makes Class III errors truly pernicious is that for a specific line, there does exist a set of $\vg$ that could properly fit the line, however this parameter combination would result in bad fits for the majority of the other lines. Class I and II errors are easy to identify since the lines will always be bad across a wide range of $\vg$.

In addition, it is difficult to determine the ensemble of lines that suffer from Class III errors, since the strength of some lines will be under-predicted while others will be over-predicted. How do we know which lines to trust? If we take an approach similar to fixing Class I and II errors, and mask out bad regions, it would be very difficult to distinguish Class III errors from \emph{truthful} lines at an incorrect $\vg$. For example, if one imagines using a cleaning algorithm as described for Class I and Class II errors, the identification of Class III errors will be sensitive to the starting guesses of $\vg$.

Because these Class III lines can still produce a significant residual spike, they tend to distort and in some cases artificially peak the posterior distribution. There is no one parameter combination that will fit all lines perfectly. Certainly there is a parameter combination that will produce a reasonable fit to 90\% of the lines to within 3 sigma residuals, but there will still be a fair number of both ``high'' and ``low'' regions. These lines need to be identified and dealt with, but how?

How then, should the best fit be determined? Surely some of the remaining lines are ``bad.'' But in this regime, determination of which lines are bad will bias the results in a certain direction.

Also, which lines should one actually fit? The stronger lines? Did the modelers themselves give more attention to the atomic constants which produce the strong lines?

To summarize, the fundamental problem with comparing high resolution optical stellar spectra to grids of synthetic stellar spectra is that no one model $\vg$ will fit all of the lines. Deciding how to model the lines is then the difficult part. Different groups have sigma-clipped lines that are obviously bad, and this does improve the procedure dramatically. In such a case, the procedure of sigma-clipping doesn't gall with me too badly. However, the more difficult problem is dealing with Class III errors properly, something that no modeler has mentioned or successfully implemented in a statistically justified manner.

In a spectrum fit, many lines do reasonably fit within the Poisson errors. For consistency, we will denote these as ``Class 0'' lines, where there is no error.

\subsection{Testing the accuracy of stellar libraries}
Gaia team members are panicking about the upcoming flood of data.

\citet{mc07} compare many spectral grids to determine their weaknesses. \citep{svt+11} discuss the range of stellar grids available for Gaia.

\citep{cbm+05} is the Coehlo grid.

\section{Instrumental effects and ``nuisance'' parameters in the limit of a perfect stellar model}

The parameter vector is comprised of both stellar parameters and calibration parameters $\vt = \{\vstar, \vN \}$. Stellar parameters include $\vstar = \{T_{\rm eff}, \log(g), [{\rm Fe}/{\rm H}], v \sin i, v_z, A_v, R_\star^2/d^2 \}$ while the data vector $\vD = \fD(\lambda)$ represents a high resolution spectrum. For demonstration purposes in this paper, we use a 51-order echelle spectrum from the \emph{TRES} spectrograph at a resolution of $R=48,000$ ($6.8\;{\rm km/s}$), which spans the full optical range ($3800-9000$\AA), though our technique is applicable to any high resolution optical/near-IR spectrum.

The pixel-to-pixel likelihood function will evaluate the residuals in some form of 
\begin{equation}
  {\rm residual} = \frac{ {\rm data} - {\rm model}}{ {\rm noise}}
\end{equation}
where the best model is defined as one that minimizes the residuals in some sense. We will discuss other goodness of fit functions soon. Each of these quantities are functions of wavelength $\lambda$, so moving forward we will refer to them as the data spectrum $\fD(\lambda)$, the model spectrum $\fM(\lambda)$, the noise spectrum $\sigma(\lambda)$, and the residual spectrum $R(\lambda)$.

FIGURE showing the high resolution, raw $R =$ 500,000 spectrum compared to the one convolved with a kernel down to $R =$ 50,000. This shows that lines are blended, but it is important to keep in mind that we do have access to the high resolution models when talking about bad pixels or bad line-species. For example, we could inject bad lines in this space.

\subsection{Processing TRES Data}
We flux calibrate TRES data, but because the reduction pipeline does not produce a ``sigma spectrum,'' we instead use the Poisson errors on each pixel. By taking the raw, still-blazed spectrum, in units of ``counts,'' we can find the Signal-to-Noise ratio by,
\begin{equation}
  S/N = \sqrt{ {\rm cts}}
\end{equation}
then the Noise-to-Signal ratio is the inverse of this
\begin{equation}
  N/S = \frac{1}{\sqrt{ {\rm cts}}}
\end{equation}
then the errors on the flux-calibrated spectrum (in ergs/s/A/cm2) is simply the flux spectrum multiplied by N/S.

\subsection{Applied to a Gaussian likelihood function}

@@

Because matrix multiplication is associative, we can sum ${\bm W(\lambda)}$ and $\vec{T}^\trans(\lambda)$ across all $\lambda$, and define
\begin{align}
  {\bm A} &= c_0^2 \sum_{\lambda} \frac{\fM^2(\lambda)}{\sigma^2(\lambda)} {\bm W}(\lambda) \\
  \vec{B} &= \sum_{\lambda} \frac{-\fM^2(\lambda) c_0^2 + \fD(\lambda) \fM (\lambda) c_0}{\sigma^2(\lambda)} \vec{T}(\lambda)\\
  g &= -\frac{1}{2} \sum_{\lambda} \frac{c_0^2 \fM^2(\lambda) - 2 c_0 \fD(\lambda) \fM(\lambda) + \fD^2(\lambda)}{\sigma^2(\lambda)} 
\end{align}
Where ${\bm A}$, $\vec{B}$, and $g$ are each a function of $\vstar$, $\vN$, and $\vD$. Now we can rewrite Equation~\ref{eqn:lnprob2} as 
\begin{align}
  \ln \bigl [p(\vstar, \vN | \vD) \bigr] &\propto - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans \vc + g\\
  p(\vt | \vD) = p(\vstar, \vN | \vD) &\propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans \vc + g \right )
  \label{eqn:lnprob_matrix}
\end{align}
Here we have the full posterior probability distribution assuming flat priors. This posterior is a function of both the stellar parameters $\vstar$ and the nuisance parameters $\vN$. Generally, we are most interested in the stellar parameters after they have been marginalized over the nuisance parameters. This marginalization encapsulates the uncertainty due to the inference on the nuisance parameters. Because this probability function is a multi-dimensional Gaussian in the higher order Chebyshev coefficients $\vc = \{c_1, c_2, \ldots, c_N \}$, we can do this marginalization analytically, and then sample from the marginalized distribution. 
\begin{equation}
  p(\vstar, c_0 | \vD) = \int p(\vstar, c_0, \vc\; | \vD) \dd \vc
\end{equation}
The analytic multi-dimensional Gaussian integral with linear term \citep{sgd+09} yields 
\begin{align}
  p(\vstar, c_0 | \vD) &\propto \int \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans \vc + g \right ) \dd \vc\\
  p(\vstar, c_0 | \vD) &\propto \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A} \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^\trans {\bm A}^{-1} \vec{B} + g \right )}
\end{align}
Where $N$ is the number of Chebyshev coefficients marginalized over, in our case $N=3$. By doing this marginalization we have reduced the dimensionality of our posterior space from $200+$ parameters ($\sim 7$ stellar parameters plus 3 Chebyshev coefficients for each of 51 echelle orders) to $\sim 60$ ($\sim 7$ stellar parameters and one $c_0$ for each of the 51 echelle orders). This enables a dramatic speedup in the MCMC sampling. For flux-calibrated spectra where the variation in $c_0$ is expected to be small (less than XX\%), we can approximate $k$ in a different manner which allows marginalization over the $c_0$ coefficients as well, reducing the dimensionality of the space to only the $\sim 7$ stellar parameters. This approach is detailed in \S\ref{sec:gaussian_simplification}.

\subsection{Including Gaussian priors on nuisance parameters}
\label{sec:priors}
Using this formalism, it is possible to incorporate prior knowledge about the degree of necessary ``re-fluxing'' through priors on the nuisance parameters. From our flux-calibration experiment, we determined that the overall scaling could range by 10\%, while the higher order corrections were less than 3\%. Because $c_0$ is a scale factor, it should be equally probable to scale up by a factor of two as it is to scale down by a factor of two. This means we should be using a log-normal prior on $c_0$
\begin{align}
  p(c_0) &= \frac{1}{\sqrt{2 \pi} \sigma_{c_0} c_0} \exp \left( -\frac{(\ln c_0)^2}{2 \sigma_{c_0}^2} \right) \\
  \ln p(c_0) &= \ln \left( \frac{1}{\sqrt{2 \pi} \sigma_{c_0} c_0} \right) - \frac{(\ln c_0)^2}{2 \sigma_{c_0}^2}
\end{align}
centered on $c_0 = 1$. Because the higher order polynomial terms are perturbations on the scaled model ($c_0 f$), we assume Gaussian priors centered about 0
\begin{equation}
  p(c_n) = \frac{1}{\sqrt{2 \pi} \sigma_{c_n}} \exp \left( - \frac{c_n^2}{2 \sigma_{c_n^2}} \right)
\end{equation}
which we can express in vector form as
\begin{equation}
  p(\vc\;) \propto \exp \left ( -\frac{1}{2} (\vc - \vec{\mu})^\trans {\bm D} (\vc - \vec{\mu}) \right )
  \label{eqn:nuisance_prior} 
\end{equation}
where 
\begin{equation}
  {\bm D} = 
  \begin{bmatrix}
    \sigma_1^{-2} & 0 & \hdots & 0 \\
    0 & \sigma_2^{-2} & \hdots & 0 \\
    \vdots & \vdots & \ddots & 0 \\
    0 & 0 & \hdots & \sigma_N^{-2} \\
  \end{bmatrix}
\end{equation}
and $\sigma_i$ represents the width of the Gaussian prior on the i-th Chebyshev coefficient, in our case $\sigma_i = \sigma_{c_n} \approx 0.03$. $\vec{\mu}$ is the mean of the Gaussian prior, in nearly all cases we expect to set this to $\vec{\mu} = \{0, 0, \ldots, 0\}$, unless we have prior knowledge that the flux-calibration is systematically warped in one direction. Now, we can write the posterior with priors as
\begin{equation}
  p(\vstar, \vN | \vD) = p(\vstar, c_0, \vc\; | \vD) \propto p( \vD | \vstar, c_0, \vc\; ) p(\vc\;) p(\vstar) p(c_0) 
\end{equation}
\begin{equation}
  p(\vstar, \vN | \vD) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans(\vstar) \vc + g(\vstar) \right )  \exp \left ( -\frac{1}{2} (\vc - \vec{\mu})^\trans {\bm D} (\vc - \vec{\mu}) \right ) p(\vstar) p(c_0)
  \label{eqn:posterior_prior}
\end{equation}
we can expand and rearrange the argument of Equation~\ref{eqn:nuisance_prior} to a similar form that is quadratic in $\vc$
\begin{align}
  -\frac{1}{2} (\vc - \vec{\mu})^\trans {\bm D} (\vc - \vec{\mu}) &= \frac{1}{2}\left ( -\vc^\trans {\bm D} \vc + \vc^\trans {\bm D} \vec{\mu} + \vec{\mu}^\trans {\bm D} \vc - \vec{\mu}^\trans {\bm D} \vec{\mu} \right )\\
  &= -\frac{1}{2} \vc^\trans {\bm D} \vc + ({\bm D} \vec{\mu})^\trans \vc - \frac{1}{2} \vec{\mu}^\trans {\bm D} \vec{\mu}
\end{align}
then we can rewrite
\begin{align}
  {\bm A}^\prime &= {\bm A} + {\bm D}\\
  \vec{B}^\prime &= \vec{B} + ({\bm D} \vec{\mu})^\trans\\
  g^\prime &= g - \frac{1}{2} \vec{\mu}^\trans {\bm D} \vec{\mu} 
\end{align}
in the case that $\vec{\mu} = \{0, 0, \ldots, 0\}$, there is only a non-zero correction to ${\bm A}$. The full posterior probability distribution is then
 \begin{equation}
   \boxed{
  p(\vstar, \vN | \vD) = p(\vstar, c_0, \vc\;) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A}^\prime \vc + \vec{B}^{\prime \trans} \vc + g^\prime \right ) p(\vstar) p(c_0) p(\vc)
}
\end{equation}
This probability function appears in the code as \texttt{lnprob\_lognormal} and has $\sim 7$ stellar dimensions and $\sim 200$ nuisance dimensions.
